Investigating Neolithic pottery use from Guijiabao in the Yanyuan Basin in southwest China using organic residue analysis

### 2023-03-13 analysis/paper/paper.Rmd
minor edits

<!--the major finding at GJB-->
The common vessel forms unearthed at Guijiabao include jars, urns, bowls, and basins. The representative ceramic styles in the Neolithic period are grayish high-tempered fine ware and reddish brown coarse ware with decorations such as cord-marked and basket-impressed patterns [@Hao2022; @Guijiabao2017a]. Some ceramic decorations present northern styles, and some ceramics with incised geometric patterns and impressed dot patterns resemble those found in western Yunnan, indicative of relationships between Guijiabao and sites in Yunnan and the southern part of the Hengduan Mountains [@Guijiabao2017a]. Plant remains reveal the cultivation of crops dominated by foxtail millet (*Setaria italica*) and broomcorn millet (*Panicum miliaceum*) based on flotation samples collected from ash pits and stratigraphic layers that can be traced back to 5000 BP [@Guijiabao2016; @Hao2022]. This corresponds to a C4-based diet based on isotopic analysis of four human bones [@Lin2022]. In addition to millet, a smaller number of rice grains (*Oryza sativa*) have also been found in ash pits, where the earliest one dates to 4700 BP [@Hao2022]. A detailed phytolith analysis for soil samples dating to the Neolithic period (5000-3700 BP) further suggests the presence of rice phytoliths throughout the temporal sequences, suggesting a mixed millet/rice cultivation prior to 5000 BP [@Huan2022].

>>>>

<!--the major finding at GJB-->
The common vessel forms unearthed at Guijiabao include jars, urns, bowls, and basins. The representative ceramic styles in the Neolithic period are grayish high-tempered fine ware and reddish brown coarse ware with decorations such as cord-marked and basket-impressed patterns [@Hao2022; @Guijiabao2017a]. Some ceramic decorations present northern styles, and some ceramics with incised geometric patterns and impressed dot patterns resemble those found in western Yunnan, indicative of relationships between Guijiabao and sites in Yunnan and the southern part of the Hengduan Mountains [@Guijiabao2017a]. Plant remains reveal the cultivation of crops dominated by foxtail millet (*Setaria italica*) and broomcorn millet (*Panicum miliaceum*) based on flotation samples collected from ash pits and stratigraphic layers that can be traced back to 5000 BP [@Guijiabao2016; @Hao2022]. This corresponds to a C4-based diet based on isotopic analysis of four human bones [@Lin2022]. In addition to millet, a smaller number of rice grains (*Oryza sativa*) have also been found in ash pits, where the earliest one dates to 4700 BP [@Hao2022]. A detailed phytolith analysis for soil samples dating to the Neolithic period (5000-3700 BP) further suggests the presence of rice phytoliths throughout the temporal sequences [@Huan2022].
